---
title: "File S1"
author: "Jullien Flynn"
date: '2017-08-01'
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is the first part of analysis including for the paper "Rates and patterns of mutation in tandem repetitive DNA in six independent lineages of Chamydomonas reinhardtii". 
This file goes through the genome-wide description of tandem repeats in 6 different strains of Chlamydomonas reinhardtii.  

First, load necessary libraries:  

```{r, message=FALSE, warning=FALSE, include=FALSE}
remove(list = ls())
library(adegenet)
library(ggplot2)
library(ape)
library(Biostrings)
library(ggbiplot)
library(corrplot)
source("http://bioconductor.org/biocLite.R")
require(Biostrings)
```
# Look at proportion of reads containing repeats 

```{r}
reads.summary <- read.delim("~/Documents/AndyLab/Chlamy/R_Analysis/data_in/reads.summary.txt", header=FALSE)
colnames(reads.summary) <- c("File", "Repreads", "Totreads", "Proportion")
plot(x=reads.summary$Totreads, y=reads.summary$Repreads, xlab="Total reads", ylab="Reads with tandem repeats")
hist(reads.summary$Proportion)
summary(lm(formula = reads.summary$Repreads ~ reads.summary$Totreads))
# p < 2.2e-16, R^2 = 0.88

# summaries:
sum(as.numeric(reads.summary$Totreads))
(mean(reads.summary$Proportion))*100
sum(as.numeric(reads.summary$Repreads))
```
# Load in and process dataset  
This dataset is the counts of all kmers found by k-Seek in the 85 MA lines analyzed. Before running k-Seek, the data have been trimmed using trimmomatic to remove low quality sequences. The kmer counts have been normalized by GC content and depth. 

```{r}
data <- read.csv('/Users/jullienflynn/Documents/AndyLab/Chlamy/R_Analysis/data_in/Chlamy_trimmed_GCcorr.compiled')
file_names <- as.vector(data$X)
rownames(data) <- file_names
data[,1] <- NULL

# match the file names to the names of the lines
Chlamy_file_info <- read.delim("~/Documents/AndyLab/Chlamy/R_Analysis/info/Chlamy_file_info.txt", header=FALSE)
Chlamy_lines <- as.vector(Chlamy_file_info[,2])
Chlamy_files <- as.vector(Chlamy_file_info[,1])
names(Chlamy_lines) <- Chlamy_files
Chlamy_lines <- as.data.frame(Chlamy_lines)

data_withnames <- merge(data, Chlamy_lines, by=0)
lines <- data_withnames[,3314]
data_withnames <- data_withnames[,-c(1,3314)]
rownames(data_withnames) <- lines
data_withnames <- as.matrix(data_withnames)

# sort the columns (kmers) by their mean abundance
kmer_means <- colMeans(data_withnames)
data_sorted <- data_withnames[,order(kmer_means, decreasing=T)]

```
## Quality check of GC correction  
Some MA lines had their kmer counts corrected strongly down for GC contents ~50%. Check to see if it appears like these MA lines appear to have been over-corrected in the counts of kmers 50%. Look at the AC repeat in particular
```{r}
# format the data and name the rows by the MA line ID
data_withnames <- merge(data, Chlamy_lines, by=0)
ncol(data_withnames)
lines <- data_withnames[,3314]
data_withnames <- data_withnames[,-c(1,3314)]
rownames(data_withnames) <- lines
data_withnames <- as.matrix(data_withnames)

# sort the columns (kmers) by their mean abundance
kmer_means <- colMeans(data_withnames)
data_sorted <- data_withnames[,order(kmer_means, decreasing=T)]

# Some MA lines were strongly corrected near 50% GC - check if this seems reasonable by the corrected copy number  
data_sorted["CC-2937_MA1", 1:5] 
data_sorted["CC-2937_MA2", 1:5]

```
# Get the abundant kmers  
## Abundant kmers within each ancestor 

```{r}
# function that takes in a matrix of copy numbers and returns the indexes of kmers that have a mean copy number across all MA lines of at least 100
get_common_kmers <- function (kmer_matrix) {
  common.kmer.indexes <- c()
  for (i in 1:ncol(kmer_matrix)) {
    if (mean(kmer_matrix[,i])>=100) {
      common.kmer.indexes <- c(common.kmer.indexes, i)
    }
  }
  names(common.kmer.indexes) <- colnames(kmer_matrix)[common.kmer.indexes]
  return (common.kmer.indexes)
}

# make a new matrix for each ancestral group  
data_1373 <- data_sorted[which (substr(rownames(data_sorted), start = 1, stop = 7) == "CC-1373"),  ]
data_1952 <- data_sorted[which (substr(rownames(data_sorted), start = 1, stop = 7) == "CC-1952"),  ]
data_2342 <- data_sorted[which (substr(rownames(data_sorted), start = 1, stop = 7) == "CC-2342"),  ]
data_2344 <- data_sorted[which (substr(rownames(data_sorted), start = 1, stop = 7) == "CC-2344"),  ]
data_2931 <- data_sorted[which (substr(rownames(data_sorted), start = 1, stop = 7) == "CC-2931"),  ]
data_2937 <- data_sorted[which (substr(rownames(data_sorted), start = 1, stop = 7) == "CC-2937"),  ]

# get the common kmer indexes for each ancestral group, and how many each has
common.kmer.indexes_1373 <- get_common_kmers(data_1373)
length(common.kmer.indexes_1373) # 28

common.kmer.indexes_1952 <- get_common_kmers(data_1952)
length(common.kmer.indexes_1952) # 31

common.kmer.indexes_2342 <- get_common_kmers(data_2342)
length(common.kmer.indexes_2342) # 31

common.kmer.indexes_2344 <- get_common_kmers(data_2344)
length(common.kmer.indexes_2344) # 26

common.kmer.indexes_2931 <- get_common_kmers(data_2931)
length(common.kmer.indexes_2931) # 29

common.kmer.indexes_2937 <- get_common_kmers(data_2937)
length(common.kmer.indexes_2937) # 29

```
## Get the kmers that have at least 1000 copies  

```{r}
get_1000_kmers <- function (kmer_matrix) {
  common.kmer.indexes <- c()
  for (i in 1:ncol(kmer_matrix)) {
    if (mean(kmer_matrix[,i])>=1000) {
      common.kmer.indexes <- c(common.kmer.indexes, i)
    }
  }
  names(common.kmer.indexes) <- colnames(kmer_matrix)[common.kmer.indexes]
  return (common.kmer.indexes)
}
kmer1000.indexes_1373 <- get_1000_kmers(data_1373)
kmer1000.indexes_1373 # length 4

kmer1000.indexes_1952 <- get_1000_kmers(data_1952)
kmer1000.indexes_1952 # length 5

kmer1000.indexes_2342 <- get_1000_kmers(data_2342)
kmer1000.indexes_2342 # length 6

kmer1000.indexes_2344 <- get_1000_kmers(data_2344)
kmer1000.indexes_2344 # length 3

kmer1000.indexes_2931 <- get_1000_kmers(data_2931)
kmer1000.indexes_2931 # length 7

kmer1000.indexes_2937 <- get_1000_kmers(data_2937)
kmer1000.indexes_2937 # length 5
```
## Get the AC repeat means  
AC is the most abundant repeat in all ancestors. 

```{r}
mean (data_1373[,which(colnames(data_1373) == "AC")])
mean (data_1952[,which(colnames(data_1952) == "AC")])
mean (data_2342[,which(colnames(data_2342) == "AC")])
mean (data_2344[,which(colnames(data_2344) == "AC")])
mean (data_2931[,which(colnames(data_2931) == "AC")])
mean (data_2937[,which(colnames(data_2937) == "AC")])
```
## Get the union and intersection of the common kmers  

```{r}
shared.common.kmers <- Reduce(intersect, list
       (names(common.kmer.indexes_1373), names(common.kmer.indexes_1952),
       names(common.kmer.indexes_2342), names(common.kmer.indexes_2344),
       names(common.kmer.indexes_2931), names(common.kmer.indexes_2937)
       ))
shared.common.kmers
length(shared.common.kmers)
#write(shared.common.kmers, file="/Users/jullienflynn/Documents/AndyLab/Chlamy/R_Analysis/data_in/chlamy_shared.common.kmers.txt", sep="\n")

common.kmer.union <- Reduce(union, list
       (names(common.kmer.indexes_1373), names(common.kmer.indexes_1952),
       names(common.kmer.indexes_2342), names(common.kmer.indexes_2344),
       names(common.kmer.indexes_2931), names(common.kmer.indexes_2937)
       ))
common.kmer.union
length(common.kmer.union)
# get the indexes of the common kmers
common.kmer.union.indexes <- Reduce(union, list
       (common.kmer.indexes_1373, common.kmer.indexes_1952,
       common.kmer.indexes_2342, common.kmer.indexes_2344,
       common.kmer.indexes_2931, common.kmer.indexes_2937
       ))

```
# PCA  
Use the union of common kmers (46), as well as the overall top 100 kmers to make the PCA.  

```{r}
get_ancestor <- function(row_number) {
  Anc <- substr(rownames(data_sorted)[row_number], start = 1, stop = 7)
  return (Anc)
}

category <- c()
for (i in 1:nrow(data_sorted)) {
  category <- c(category, get_ancestor(i))
}
pca.data <- prcomp(data_sorted[, common.kmer.union.indexes], scale.=T)

# using the 46 union kmers
g <- ggbiplot(pca.data, obs.scale = 1, var.scale = 1, groups = category, var.axes = F)
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)

# what are these 4 samples that are separated from the rest of their group?

MA_names <- rownames(data_2342)
pca.data_2342 <- prcomp(data_2342[, common.kmer.indexes_2342], scale.=T)

g <- ggbiplot(pca.data_2342, obs.scale = 1, var.scale = 1, groups = MA_names, var.axes = F)
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)
data_2342[,1:10]
# likely MA1 and MA5 are outliters because of poly-C repeat, MA3 and MA4 because of AGGC repeat

# what about the 1952 sample?
MA_names <- rownames(data_1952)
pca.data_1952 <- prcomp(data_1952[, common.kmer.indexes_1952], scale.=T)

g <- ggbiplot(pca.data_1952, obs.scale = 1, var.scale = 1, groups = MA_names, var.axes = F)
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)
data_1952[,1:10]
# CC1952-MA9 is an outlier probably mainly driven by lower number of telomere repeat

# Now using PCs 2-3  

g <- ggbiplot(pca.data, obs.scale = 1, var.scale = 1, groups = category, var.axes = F, choices=2:3)
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)

g <- ggbiplot(pca.data, obs.scale = 1, var.scale = 1, groups = category, var.axes = F, choices=c(1,3))
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)

# Using the overall top 100 kmers

pca.data <- prcomp(data_sorted[, 1:100], scale.=T)

g <- ggbiplot(pca.data, obs.scale = 1, var.scale = 1, groups = category, var.axes = F)
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)

g <- ggbiplot(pca.data, obs.scale = 1, var.scale = 1, groups = category, var.axes = F, choices=2:3)
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)

g <- ggbiplot(pca.data, obs.scale = 1, var.scale = 1, groups = category, var.axes = F, choices=c(1,3))
g <- g + scale_color_discrete(name='')
g <- g + theme(legend.direction='horizontal', legend.position='top')
print(g)

```
# Make Figure 1  
Choose a few kmers to illustrate  

```{r}

Fig1_kmers <- c("AGGCGG", "AAC", "ACACGC", "ACCGGC", "AGGCGC")
Ancestors <- c(rep("CC-1373", times=5), rep("CC-1952", times=5), rep("CC-2342", times=5), rep("CC-2344", times=5), rep("CC-2931", times=5), rep("CC-2937", times=5))

#Ancestors

# make a data frame
means_1373 <- c()
means_1952 <- c()
means_2342 <- c()
means_2344 <- c()
means_2931 <- c()
means_2937 <- c()
for (i in 1:length(Fig1_kmers)) {
    means_1373 <- c(means_1373, mean(data_1373[,Fig1_kmers[i]]))
    means_1952 <- c(means_1952, mean(data_1952[,Fig1_kmers[i]]))
    means_2342 <- c(means_2342, mean(data_2342[,Fig1_kmers[i]]))
    means_2344 <- c(means_2344, mean(data_2344[,Fig1_kmers[i]]))
    means_2931 <- c(means_2931, mean(data_2931[,Fig1_kmers[i]]))
    means_2937 <- c(means_2937, mean(data_2937[,Fig1_kmers[i]]))
}
means_fig1 <- c(means_1373, means_1952, means_2342, means_2344, means_2931, means_2937)

Fig1_df <- data.frame(Ancestors, Fig1_kmers, means_fig1)

ggplot(Fig1_df, aes(fill=Ancestors, x=Fig1_kmers, y=means_fig1)) + 
  geom_bar(position="dodge", stat="identity") + 
  ylab("Mean copy number") +
  scale_fill_brewer(palette="Set1") +
  theme_bw(base_size = 14) + 
  theme(axis.title.x=element_blank())

```
# Length and GC distribution of kmers   

``` {r} 
# function to calculate length of string
Calc_kmer_length <- function (kmer_string) {
  return (nchar(kmer_string))
}

# function to calculate GC content of a kmer
Calc_GC_content <- function (kmer_string) {
  string <- BString(kmer_string)
  gc.content <- c(letterFrequency(string, letters=c("CG"), as.prob=F)/(length(string)))
  return(gc.content)
}

common.kmer.union.lengths <- sapply(common.kmer.union, Calc_kmer_length)
common.kmer.union.gc <- sapply(common.kmer.union, Calc_GC_content)

hist(common.kmer.union.lengths, breaks=11)
hist(common.kmer.union.gc, breaks=10)

# 6mers are the most abundant
length(which(common.kmer.union.lengths == 6))

```

# Make Figure 2

```{r}
qplot(x=common.kmer.union.lengths, geom="histogram", bins=15) + xlab("Kmer unit length") + 
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.title = element_text(size=13))

qplot(x=common.kmer.union.gc, geom="histogram", bins=15) + scale_x_reverse() + xlab("GC of kmer") + 
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.title = element_text(size=13))

# Make an extra table
table2 <- data.frame(common.kmer.union, common.kmer.union.lengths, common.kmer.union.gc)

#write.table(table2, file = "/Users/jullienflynn/Documents/AndyLab/Chlamy/ManuscriptFigures/Table2.txt", row.names = F, quote=F)

```
# Interesting pattern of GC content  
Maybe some repeats are coming from the chloroplast given their GC contents and the previous finding that the chloroplast DOES contain repeats.  

```{r}
mean(common.kmer.union.gc)

hist (sapply (names(which(common.kmer.union.lengths == 6)), Calc_GC_content), main="6-mer GC", xlab="GC")
hist (sapply (names(which(common.kmer.union.lengths != 6)), Calc_GC_content), main="non-6mer GC", xlab="GC")

# all the longer kmers >=10 bp are GC < 0.35
hist (sapply (names(which(common.kmer.union.lengths >= 10)), Calc_GC_content), main="10+ mer GC", xlab="GC")

hist (sapply (names(which(common.kmer.union.lengths <= 9)), Calc_GC_content), main="<9 mer GC", xlab="GC")
length(which(common.kmer.union.lengths <= 9))

mean(sapply (names(which(common.kmer.union.lengths >= 10)), Calc_GC_content))
# average GC is 17.4 !!
max(sapply (names(which(common.kmer.union.lengths >= 10)), Calc_GC_content))

mean(sapply (names(which(common.kmer.union.lengths <= 9)), Calc_GC_content))

# make Figure 2c

assign_group <- function(length){
  if (length >= 10) {
    group <- 1
  }
  else {
    group <- 2
  }
  return(group)
}

group_assignments <- sapply (common.kmer.union.lengths, assign_group)
group_assignments

cp_plot <- data.frame(lengths=common.kmer.union.lengths, gc=common.kmer.union.gc, grp=as.factor(group_assignments))

ggplot (data = cp_plot, aes(y = cp_plot$gc, x = cp_plot$lengths, color = cp_plot$grp )) +
  geom_jitter(size=3, alpha=0.5) +
  xlab("Kmer unit length") + 
  ylab("GC content") +
  theme_bw() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=14), legend.position="none")

```
# Differences in the abundance of the telomere repeat  

```{r}
temp <- which(colnames (data_1373)=="AAAACCCT")
telomere_abun_1373 <- data.frame(data_1373[,temp], "CC-1373")
colnames(telomere_abun_1373) <- c("copynumber", "Ancestor")
telomere_mean_1373 <- mean(data_1373[,temp]) 

temp <- which(colnames (data_1952)=="AAAACCCT")
telomere_abun_1952 <- data.frame(data_1952[,temp], "CC-1952")
colnames(telomere_abun_1952) <- c("copynumber", "Ancestor")
telomere_mean_1952 <- mean(data_1952[,temp]) 

temp <- which(colnames (data_2342)=="AAAACCCT")
telomere_abun_2342 <- data.frame(data_2342[,temp], "CC-2342")
colnames(telomere_abun_2342) <- c("copynumber", "Ancestor")
telomere_mean_2342 <- mean(data_2342[,temp])

temp <- which(colnames (data_2344)=="AAAACCCT")
telomere_abun_2344 <- data.frame(data_2344[,temp], "CC-2344")
colnames(telomere_abun_2344) <- c("copynumber", "Ancestor")
telomere_mean_2344 <- mean(data_2344[,temp])

temp <- which(colnames (data_2931)=="AAAACCCT")
telomere_abun_2931 <- data.frame(data_2931[,temp], "CC-2931")
colnames(telomere_abun_2931) <- c("copynumber", "Ancestor")
telomere_mean_2931 <- mean(data_2931[,temp])

temp <- which(colnames (data_2937)=="AAAACCCT")
telomere_abun_2937 <- data.frame(data_2937[,temp], "CC-2937")
colnames(telomere_abun_2937) <- c("copynumber", "Ancestor")
telomere_mean_2937 <- mean(data_2937[,temp])

telomere_abun_df <- rbind(telomere_abun_1373, telomere_abun_1952, telomere_abun_2342,
                          telomere_abun_2344, telomere_abun_2931, telomere_abun_2937)

boxplot(telomere_abun_df[,1] ~ telomere_abun_df[,2], data=telomere_abun_df, las=2, ylab = "Telomere copy number")
# seems like CC-2344 has higher telomere copy numbers

telomere_model <- aov(copynumber ~ Ancestor, data = telomere_abun_df)
summary(telomere_model)
# anova indicates significant difference
posthoc_telomere <- TukeyHSD(x=telomere_model, conf.level = 0.95)
posthoc_telomere
help("TukeyHSD")
hist(telomere_abun_1373$copynumber)
hist(telomere_abun_1952$copynumber)
hist(telomere_abun_2342$copynumber)
hist(telomere_abun_2344$copynumber)
hist(telomere_abun_2931$copynumber)
hist(telomere_abun_2937$copynumber)

get_residuals <- function (abun_mx) {
  x <- abun_mx$copynumber - mean(abun_mx$copynumber)
  return(x)
}
residuals_1373 <- get_residuals(telomere_abun_1373)
residuals_1952 <- get_residuals(telomere_abun_1952)
residuals_2342 <- get_residuals(telomere_abun_2342)
residuals_2344 <- get_residuals(telomere_abun_2344)
residuals_2931 <- get_residuals(telomere_abun_2931)
residuals_2937 <- get_residuals(telomere_abun_2937)
residuals_all <- c(residuals_1373, residuals_1952, residuals_2342, residuals_2344, residuals_2931, residuals_2937)
hist(residuals_all)
# great, the residuals are approximately normally distributed, parametric test is fine to do

```
# Kmers that are polymorphic by presence/absence   

```{r}

# For Figure 1B
mean(data_1373[,"AAATACCTTACGGGAATAT"])
mean(data_1952[,"AAATACCTTACGGGAATAT"])
mean(data_2342[,"AAATACCTTACGGGAATAT"])
mean(data_2344[,"AAATACCTTACGGGAATAT"])
mean(data_2931[,"AAATACCTTACGGGAATAT"])
mean(data_2937[,"AAATACCTTACGGGAATAT"])

mean(data_1373[,"AAATAACAAT"])
mean(data_1952[,"AAATAACAAT"])
mean(data_2342[,"AAATAACAAT"])
mean(data_2344[,"AAATAACAAT"])
mean(data_2931[,"AAATAACAAT"])
mean(data_2937[,"AAATAACAAT"])

mean(data_1373[,"ATC"])
mean(data_1952[,"ATC"])
mean(data_2342[,"ATC"])
mean(data_2344[,"ATC"])
mean(data_2931[,"ATC"])
mean(data_2937[,"ATC"])

mean(data_1373[,"AAATACTTACGGGAATAT"])
mean(data_1952[,"AAATACTTACGGGAATAT"])    
mean(data_2342[,"AAATACTTACGGGAATAT"])
mean(data_2344[,"AAATACTTACGGGAATAT"])
mean(data_2931[,"AAATACTTACGGGAATAT"])
mean(data_2937[,"AAATACTTACGGGAATAT"])

mean(data_1373[,"ACCCGG"])
mean(data_1952[,"ACCCGG"])
mean(data_2342[,"ACCCGG"])
mean(data_2344[,"ACCCGG"])
mean(data_2931[,"ACCCGG"])
mean(data_2937[,"ACCCGG"])

mean(data_1373[,"AAATTTATATACTCC"])
mean(data_1952[,"AAATTTATATACTCC"])
mean(data_2342[,"AAATTTATATACTCC"])
mean(data_2344[,"AAATTTATATACTCC"])
mean(data_2931[,"AAATTTATATACTCC"])
mean(data_2937[,"AAATTTATATACTCC"])

mean(data_1373[,"AAAGAAGTATATAAAT"])
mean(data_1952[,"AAAGAAGTATATAAAT"])
mean(data_2342[,"AAAGAAGTATATAAAT"])
mean(data_2344[,"AAAGAAGTATATAAAT"])
mean(data_2931[,"AAAGAAGTATATAAAT"])
mean(data_2937[,"AAAGAAGTATATAAAT"])

mean(data_1373[,"AAAATATATAAATATAGCT"])
mean(data_1952[,"AAAATATATAAATATAGCT"])
mean(data_2342[,"AAAATATATAAATATAGCT"])
mean(data_2344[,"AAAATATATAAATATAGCT"])
mean(data_2931[,"AAAATATATAAATATAGCT"])
mean(data_2937[,"AAAATATATAAATATAGCT"])

mean(data_1373[,"AATAGATAATAT"])
mean(data_1952[,"AATAGATAATAT"])
mean(data_2342[,"AATAGATAATAT"])
mean(data_2344[,"AATAGATAATAT"])
mean(data_2931[,"AATAGATAATAT"])
mean(data_2937[,"AATAGATAATAT"])

mean(data_1373[,"AAATAGCAGTATAT"])
mean(data_1952[,"AAATAGCAGTATAT"])
mean(data_2342[,"AAATAGCAGTATAT"])
mean(data_2344[,"AAATAGCAGTATAT"])
mean(data_2931[,"AAATAGCAGTATAT"])
mean(data_2937[,"AAATAGCAGTATAT"])

mean(data_1373[,"AAAAAGGTAAATGTATTTAT"])
mean(data_1952[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2342[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2344[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2931[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2937[,"AAAAAGGTAAATGTATTTAT"])

mean(data_1373[,"AAAAAGGTAAATGTATTTAT"])
mean(data_1952[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2342[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2344[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2931[,"AAAAAGGTAAATGTATTTAT"])
mean(data_2937[,"AAAAAGGTAAATGTATTTAT"])

mean(data_1373[,"AGGGCC"])
mean(data_1952[,"AGGGCC"])
mean(data_2342[,"AGGGCC"])
mean(data_2344[,"AGGGCC"])
mean(data_2931[,"AGGGCC"])
mean(data_2937[,"AGGGCC"])

mean(data_1373[,"AGAGGC"])
mean(data_1952[,"AGAGGC"])
mean(data_2342[,"AGAGGC"])
mean(data_2344[,"AGAGGC"])
mean(data_2931[,"AGAGGC"])
mean(data_2937[,"AGAGGC"])

mean(data_1373[,"AAAATAAAGTGT"])
mean(data_1952[,"AAAATAAAGTGT"])
mean(data_2342[,"AAAATAAAGTGT"])
mean(data_2344[,"AAAATAAAGTGT"])
mean(data_2931[,"AAAATAAAGTGT"])
mean(data_2937[,"AAAATAAAGTGT"])

mean(data_1373[,"ACTGGGC"])
mean(data_1952[,"ACTGGGC"])
mean(data_2342[,"ACTGGGC"])
mean(data_2344[,"ACTGGGC"])
mean(data_2931[,"ACTGGGC"])
mean(data_2937[,"ACTGGGC"])

mean(data_1373[,"AAACAC"])
mean(data_1952[,"AAACAC"])
mean(data_2342[,"AAACAC"])
mean(data_2344[,"AAACAC"])
mean(data_2931[,"AAACAC"])
mean(data_2937[,"AAACAC"])

mean(data_1373[,"AAAC"])
mean(data_1952[,"AAAC"])
mean(data_2342[,"AAAC"])
mean(data_2344[,"AAAC"])
mean(data_2931[,"AAAC"])
mean(data_2937[,"AAAC"])

```

# Get total kmer content
For each ancestral lineage, get the absolute amount of kmer content in kb. Use all the kmers that have at least 2 copies (normalized) in all the MA lines  

```{r}
# get kmers that will be used for the overall genome-wide description of kmers in the 6 ancestors
# have at least 2 normalized copies in all MA lines from the a given ancestor

get_legit_kmer_indexes <- function (kmer_mx) {
  indexes <- c()
  for (i in 1:ncol(kmer_mx)) { ## go through each column of the data
    p <- kmer_mx[,i] >= 2 ## at least 2 copies
    if (sum(p)==nrow(kmer_mx)) { # if all samples have at least 2 copies
      indexes <- c(indexes, i)
    }
  }
  return(indexes)
}
#get matrix with absolute kmer content (multiplied by length of kmer)

get_abskmer_mx <- function (kmer_mx, indexes) {
  kmers.abs <- matrix(nrow=nrow(kmer_mx), ncol=max(indexes))
  for (i in indexes) {
    for (j in 1:nrow(kmer_mx)) {
    kmers.abs[j,i] <- kmer_mx[j,i] * (nchar(colnames(kmer_mx)[i], type="chars")) 
    }
  }
  kmers.abs <- kmers.abs[,colSums(is.na(kmers.abs)) != nrow(kmers.abs)] # get rid of columns that are all NAs
  return(kmers.abs)
}

# get total kmer content per MA line
get_total_kmercontent <- function (abs_mx) {
  total.kmers.abs <- c()
  for (i in 1:nrow(abs_mx)){
    total.kmers.abs[i] <- sum(abs_mx[i,])/10^3
  }
  names(total.kmers.abs) <- rownames(abs_mx)
  return (total.kmers.abs)
}


legit.kmer.indexes_1373 <- get_legit_kmer_indexes(data_1373)
length(legit.kmer.indexes_1373)
abs.mx_1373 <- get_abskmer_mx(data_1373, legit.kmer.indexes_1373)
total.kmercontent_1373 <- get_total_kmercontent(abs.mx_1373)
mean(total.kmercontent_1373)
total.kmercontent_1373 # note: this is in kb


legit.kmer.indexes_1952 <- get_legit_kmer_indexes(data_1952)
length(legit.kmer.indexes_1952) 
abs.mx_1952 <- get_abskmer_mx(data_1952, legit.kmer.indexes_1952)
total.kmercontent_1952 <- get_total_kmercontent(abs.mx_1952)
mean(total.kmercontent_1952) # note: this is in kb

legit.kmer.indexes_2342 <- get_legit_kmer_indexes(data_2342)
length(legit.kmer.indexes_2342) 
abs.mx_2342 <- get_abskmer_mx(data_2342, legit.kmer.indexes_2342)
total.kmercontent_2342 <- get_total_kmercontent(abs.mx_2342)
mean(total.kmercontent_2342) # note: this is in kb

legit.kmer.indexes_2344 <- get_legit_kmer_indexes(data_2344)
length(legit.kmer.indexes_2344) 
abs.mx_2344 <- get_abskmer_mx(data_2344, legit.kmer.indexes_2344)
total.kmercontent_2344 <- get_total_kmercontent(abs.mx_2344)
mean(total.kmercontent_2344)

legit.kmer.indexes_2931 <- get_legit_kmer_indexes(data_2931)
length(legit.kmer.indexes_2931) 
abs.mx_2931 <- get_abskmer_mx(data_2931, legit.kmer.indexes_2931)
total.kmercontent_2931 <- get_total_kmercontent(abs.mx_2931)
mean(total.kmercontent_2931)

legit.kmer.indexes_2937 <- get_legit_kmer_indexes(data_2937)
length(legit.kmer.indexes_2937) 
abs.mx_2937 <- get_abskmer_mx(data_2937, legit.kmer.indexes_2937)
total.kmercontent_2937 <- get_total_kmercontent(abs.mx_2937)
mean(total.kmercontent_2937)

# Now, make a boxplot of this.

category <- c( 
  rep ("CC-1373", times = nrow(data_1373)),
  rep ("CC-1952", times = nrow(data_1952)),
  rep ("CC-2342", times = nrow(data_2342)),
  rep ("CC-2344", times = nrow(data_2344)),
  rep ("CC-2931", times = nrow(data_2931)),
  rep ("CC-2937", times = nrow(data_2937))
  )

total.kmercontent_all <- c(total.kmercontent_1373, total.kmercontent_1952, total.kmercontent_2342, total.kmercontent_2344, total.kmercontent_2931, total.kmercontent_2937)
length(total.kmercontent_all)

TotalAbs_df <- data.frame (total.kmercontent_all, category)
par(mar=c(5,5,3,1))
boxplot (TotalAbs_df[,1] ~ TotalAbs_df[,2], data=TotalAbs_df, las=2, cex.axis=0.8, ylab="Total tandem repeat content (kb)", Main = "Per MA line absolute total mutation rate", col="deepskyblue3")
```

##Do the kmers that have similar patterns of correlation in multiple MA ancestors have similar copy number?  
i.e. could similarity in mean copy number explain the similarity in correlation patterns.  
ANSWER: NO.

```{r}
shared_corr_kmers <- c("AGGC", "AGGCGG", "ACGC", "ACGCC", "ACGCCC", "ACCGCC", "ACCCGC", "ACACGC", "ACAGGC", "ACAGGG", "ACCAGC", "AGGCGC")

data_1373[,shared_corr_kmers]
shared_corr_kmers_1373_means <- colMeans(data_1373[,shared_corr_kmers])
shared_corr_kmers_1373_means
shared_corr_kmers_1952_means <- colMeans(data_1952[,shared_corr_kmers])
shared_corr_kmers_2342_means <- colMeans(data_2342[,shared_corr_kmers])
shared_corr_kmers_2344_means <- colMeans(data_2344[,shared_corr_kmers])
shared_corr_kmers_2931_means <- colMeans(data_2931[,shared_corr_kmers])
shared_corr_kmers_2937_means <- colMeans(data_2937[,shared_corr_kmers])

shared_corr_kmers_df <- data.frame(shared_corr_kmers_1373_means, shared_corr_kmers_1952_means, shared_corr_kmers_2342_means, shared_corr_kmers_2344_means, shared_corr_kmers_2931_means, shared_corr_kmers_2937_means)
colnames(shared_corr_kmers_df) <- c("1373", "1952", "2342", "2344", "2931", "2937")

plot(shared_corr_kmers_df["AGGC",])
plot(shared_corr_kmers_df)

shared_corr_kmers_df
shared_corr_kmers_corr <- cor(shared_corr_kmers_df)
shared_corr_kmers_corr

# the means of these kmers are not the most correlated.