---
title: "File S2"
author: "Jullien Flynn"
date: '2018-01-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jullienflynn/Documents/AndyLab/Chlamy/R_Analysis/data_in/")
```
**This file details the analysis of paired-end reads to determine interspersion between repeats (when there are repeats on both mate pairs).**  

Specifically, kseek was run separately for R1 and R2.  

Next, Perl scripts were used:  
```pl rep_mate.pl read_1.rep ```  
Followed by:  
```pl mp.compile.updated.pl read_1.rep.mp read_2.rep.mp outfile```  

*Note*:  one MA line per ancestor was randomly chosen to be representative, so 6 MA lines in total were used.  

**Import required packages** 

```{r, include=FALSE}
library(ggplot2)
source("http://bioconductor.org/biocLite.R")
require(Biostrings)
```

# Import and clean up data

``` {r}
rm(list=ls())
CC_2931_MA1.rep <- read.delim("ERR968392.out.rep.mpx")
CC_2937_MA4.rep <- read.delim("ERR968409.out.rep.mpx")
CC_1952_MA2.rep <- read.delim("ERR968352.out.rep.mpx")
CC_2342_MA7.rep  <- read.delim("ERR968369.out.rep.mpx")
CC_2344_MA15.rep <- read.delim("ERR968391.out.rep.mpx")
CC_1373_MA7.rep <- read.delim("ERR968344.out.rep.mpx")

# import the names of the common kmers you are interested in.
# already have the common kmer union in my environment, but can import in file if not in environment
common.kmer.union <- read.table("../../chlamy_union_kmers.txt")
common.kmer.union <- as.vector(common.kmer.union$V1)

#head(CC_1373_MA7.rep)
#ncol(CC_1373_MA7.rep)

process_rep_files <- function(inputrep) {
  kmer_names <- inputrep[,1]
  inputrep$X <- NULL
  rownames(inputrep) <- kmer_names
  inputrep <- as.matrix(inputrep)
  return(inputrep)
}
CC_2931_MA1.rep <- process_rep_files(CC_2931_MA1.rep)
CC_2937_MA4.rep <- process_rep_files(CC_2937_MA4.rep)
CC_1952_MA2.rep <- process_rep_files(CC_1952_MA2.rep)
CC_2342_MA7.rep <- process_rep_files(CC_2342_MA7.rep)
CC_2344_MA15.rep <- process_rep_files(CC_2344_MA15.rep)
CC_1373_MA7.rep <- process_rep_files(CC_1373_MA7.rep)


```
 
Important analysis steps:
1) Take the mpx matrix (generated from the perl scrips) and look at the kmers of interest (the common union kmers)  
2) Make a pairwise matrix by adding the two triangles of the matrix, because identity of R1 vs R2 is not important here.
3) Calculate the interspersion metric, by bringing in data generated from external scripts. A metric is calculated for each pair of kmers i and j. Shell script `run_get_ni,nj.sh` is used to calculate values required for the interspersion metric. Python script `extract.py` is used to get the reads containing specific kmers.  


```{r}
get_intsp_metric.mx <- function(totalreadsfile, inputrep) {
  interspersion.matrix <- matrix(ncol=ncol(inputrep), nrow=nrow(inputrep))
  interspersion.matrix <- inputrep[colnames(inputrep) %in% common.kmer.union, rownames(inputrep) %in% common.kmer.union] 

  pairwise.matrix <- matrix(ncol=ncol(interspersion.matrix), nrow=nrow(interspersion.matrix))

  rownames(pairwise.matrix) <- rownames(interspersion.matrix)
  colnames(pairwise.matrix) <- colnames(interspersion.matrix)

  for (i in 1:nrow(interspersion.matrix)) {
    pairwise.matrix[i,i] <- interspersion.matrix[i,i]
    for (j in (i+1):nrow(interspersion.matrix)) {
      if (j <= nrow(interspersion.matrix)) {
        sum <- interspersion.matrix[i,j] + interspersion.matrix[j,i]
        pairwise.matrix[i,j] <- sum
      }
    }
  }

  total.reads <- read.table(totalreadsfile, quote="\"", comment.char="")
  read.counts <- as.vector(total.reads[,1])
  names(read.counts) <- common.kmer.union
  read.counts_temp <- read.counts[names(read.counts) %in% colnames(pairwise.matrix)]
  read.counts <- read.counts_temp[order(match(names(read.counts_temp), colnames(pairwise.matrix)))]
  names(read.counts) == colnames(pairwise.matrix)

  interspersion.metric.matrix <- matrix(ncol=ncol(pairwise.matrix), nrow=nrow(pairwise.matrix))
  rownames(interspersion.metric.matrix) <- rownames(pairwise.matrix)
  colnames(interspersion.metric.matrix) <- colnames(pairwise.matrix)

  for (i in 1:ncol(pairwise.matrix)) {
    for (j in i:ncol(pairwise.matrix)) { # i will equal j, for the same kmer with itself
      ni <- as.numeric(read.counts[i])
      nj <- as.numeric(read.counts[j])
      nij <- pairwise.matrix[i,j]
      interspersion.metric.matrix[i,j] <- nij / (sqrt(ni*nj))
    }
  }
return(interspersion.metric.matrix)
}
```
Using this function, generate interspersion metric matrices for all of the ancestors.  
```{r}

interspersion.metric.matrix_1373 <- get_intsp_metric.mx("ERR968344.total.reads.txt", CC_1373_MA7.rep)
interspersion.metric.matrix_1952 <- get_intsp_metric.mx("ERR968352.total.reads.txt", CC_1952_MA2.rep)
interspersion.metric.matrix_2342 <- get_intsp_metric.mx("ERR968369.total.reads.txt", CC_2342_MA7.rep)
interspersion.metric.matrix_2344 <- get_intsp_metric.mx("ERR968391.total.reads.txt", CC_2344_MA15.rep)
interspersion.metric.matrix_2931 <- get_intsp_metric.mx("ERR968392.total.reads.txt", CC_2931_MA1.rep)
interspersion.metric.matrix_2937 <- get_intsp_metric.mx("ERR968409.total.reads.txt", CC_2937_MA4.rep)

```
Make another function that calculates the percent of reads from a kmer that are interspersed with other kmers.
```{r}
get_percent_interspersed <- function(totalreadsfile, inputrep, kmername) {
  col_number <- which(colnames(inputrep) == kmername)
  sum_column <- sum(inputrep[,col_number])
  sum_row <- sum(inputrep[col_number,]) - inputrep[col_number,col_number]
  total <- sum_column + sum_row
  total.reads <-  read.table(totalreadsfile, quote="\"", comment.char="")
  read.counts <- as.vector(total.reads[,1])
  names(read.counts) <- common.kmer.union
  read.count.kmer <- read.counts[kmername]
  return (total/read.count.kmer)
}

which(colnames(CC_2931_MA1.rep) == "AC")
total.reads <-  read.table("ERR968392.total.reads.txt", quote="\"", comment.char="")
read.counts <- as.vector(total.reads[,1])
names(read.counts) <- common.kmer.union
read.count.kmer <- read.counts["AC"]

which(colnames(CC_2937_MA4.rep) == "AGGGCC")
total.reads <-  read.table("ERR968392.total.reads.txt", quote="\"", comment.char="")
read.counts <- as.vector(total.reads[,1])
names(read.counts) <- common.kmer.union
read.count.kmer <- read.counts["AC"]

get_percent_interspersed("ERR968392.total.reads.txt", CC_2931_MA1.rep, "AAAACCCT")


# function that loops through all kmers and returns a vector
get_percent_summary <- function(totalreadsfile, inputrep) {
  kmers <- colnames(inputrep)[colnames(inputrep) %in% common.kmer.union]
  result <- c()
  for (i in 1:length(kmers)) {
    result <- c(result, get_percent_interspersed(totalreadsfile, inputrep, kmers[i])) 
  }
  return(result)
}


int.summary_1373 <- get_percent_summary("ERR968344.total.reads.txt", CC_1373_MA7.rep)
int.summary_1952 <- get_percent_summary("ERR968352.total.reads.txt", CC_1952_MA2.rep)
int.summary_2342 <- get_percent_summary("ERR968369.total.reads.txt", CC_2342_MA7.rep)
int.summary_2344 <- get_percent_summary("ERR968391.total.reads.txt", CC_2344_MA15.rep)
int.summary_2931 <- get_percent_summary("ERR968392.total.reads.txt", CC_2931_MA1.rep)
int.summary_2937 <- get_percent_summary("ERR968409.total.reads.txt", CC_2937_MA4.rep)

par(mfrow=c(2,3))
barplot(int.summary_1373, names.arg=names(int.summary_1373), las=2, cex.names=0.4, cex.axis=0.8, main="CC-1373", ylab="fraction of reads")
barplot(int.summary_1952, names.arg=names(int.summary_1952), las=2, cex.names=0.4, cex.axis=0.8, main="CC-1952", ylab="fraction of reads")
barplot(int.summary_2342, names.arg=names(int.summary_2342), las=2, cex.names=0.4, cex.axis=0.8, main="CC-2342", ylab="fraction of reads")
barplot(int.summary_2344, names.arg=names(int.summary_2344), las=2, cex.names=0.4, cex.axis=0.8, main="CC-2344", ylab="fraction of reads")
barplot(int.summary_2931, names.arg=names(int.summary_2931), las=2, cex.names=0.4, cex.axis=0.8, main="CC-2931", ylab="fraction of reads")
barplot(int.summary_2937, names.arg=names(int.summary_2937), las=2, cex.names=0.4, cex.axis=0.8, main="CC-2937", ylab="fraction of reads")

mean(int.summary_1373)
mean(int.summary_1952)
mean(int.summary_2342)
mean(int.summary_2344)
mean(int.summary_2931)
mean(int.summary_2937)

```
# How many, and what percent of chloroplast-like reads are interspersed with the telomere repeat? 

```{r}

count_int_reads_telomere <- function(totalreadsfile, inputrep, kmer) {
  interspersion.matrix <- matrix(ncol=ncol(inputrep), nrow=nrow(inputrep))
  interspersion.matrix <- inputrep[colnames(inputrep) %in% common.kmer.union, rownames(inputrep) %in% common.kmer.union] 
  
  pairwise.matrix <- matrix(ncol=ncol(interspersion.matrix), nrow=nrow(interspersion.matrix))
  
  rownames(pairwise.matrix) <- rownames(interspersion.matrix)
  colnames(pairwise.matrix) <- colnames(interspersion.matrix)
  
  for (i in 1:nrow(interspersion.matrix)) {
    pairwise.matrix[i,i] <- interspersion.matrix[i,i]
    for (j in (i+1):nrow(interspersion.matrix)) {
      if (j <= nrow(interspersion.matrix)) {
        sum <- interspersion.matrix[i,j] + interspersion.matrix[j,i]
        pairwise.matrix[i,j] <- sum
      }
    }
  }
  
  total.reads <- read.table(totalreadsfile, quote="\"", comment.char="")
  read.counts <- as.vector(total.reads[,1])
  names(read.counts) <- common.kmer.union
  read.counts_temp <- read.counts[names(read.counts) %in% colnames(pairwise.matrix)]
  read.counts <- read.counts_temp[order(match(names(read.counts_temp), colnames(pairwise.matrix)))]
  names(read.counts) == colnames(pairwise.matrix)
  
  res_1 <- pairwise.matrix[kmer, "AAAACCCT"]
  res_2 <- pairwise.matrix["AAAACCCT", kmer]
  temp <- c(res_1, res_2)
  res_12 <- temp[!is.na(temp)]
  res_3 <- res_12/read.counts[kmer]
  res_final <- c(res_12, res_3)
  print(res_final)
}

count_int_reads_telomere("ERR968344.total.reads.txt", CC_1373_MA7.rep, "AAATAACAAT")
count_int_reads_telomere("ERR968344.total.reads.txt", CC_1373_MA7.rep, "AAATACCTTACGGGAATAT")
count_int_reads_telomere("ERR968344.total.reads.txt", CC_1373_MA7.rep, "AAATACTTACGGGAATAT")

cp_kmers_1373 <- c("AAATACCTTACGGGAATAT", "AAATAACAAT", "AAATACTTACGGGAATAT")
for (i in 1:length(cp_kmers_1373)) {
  count_int_reads_telomere("ERR968344.total.reads.txt", CC_1373_MA7.rep, cp_kmers_1373[i])
}

cp_kmers_1952 <- c("AATAGATAATAT", "AAATACTTACGGGAATAT", "AAAATATATAAATATAGCT", "AAAGAAGTATATAAAT", "AAATTTATATACTCC", "AAATACCTTACGGGAATAT")

for (i in 1:length(cp_kmers_1952)) {
  count_int_reads_telomere("ERR968352.total.reads.txt", CC_1952_MA2.rep, cp_kmers_1952[i])
}

cp_kmers_2342 <- c("AAAAAGGTAAATGTATTTAT")
for (i in 1:length(cp_kmers_2342)) {
  count_int_reads_telomere("ERR968369.total.reads.txt", CC_2342_MA7.rep, cp_kmers_2342[i])
}

cp_kmers_2344 <- c("AAAATAAAGTGT", "AAATACCTTACGGGAATAT", "AAATTTATATACTCC")
for (i in 1:length(cp_kmers_2344)) {
  count_int_reads_telomere("ERR968391.total.reads.txt", CC_2344_MA15.rep, cp_kmers_2344[i])
}

cp_kmers_2931 <- c("AAAAAGGTAAATGTATTTAT", "AAATAGCAGTATAT", "AAAATAAAGTGT")
for (i in 1:length(cp_kmers_2931)) {
  count_int_reads_telomere("ERR968392.total.reads.txt", CC_2931_MA1.rep, cp_kmers_2931[i])
}

cp_kmers_2937 <- c("AAATACTTACGGGAATAT", "AAATACCTTACGGGAATAT")
for (i in 1:length(cp_kmers_2937)) {
  count_int_reads_telomere("ERR968409.total.reads.txt", CC_2937_MA4.rep, cp_kmers_2937[i])
}


```
# Get kmers that are interspersed with each other
Kmers that have an interspersion metric > 0 are interspersed with each other.  


``` {r}
# function that takes a kmer sequence and an interspersion matrix, and returns the kmers that are interspersed with the given input kmer (along with the corresponding interspersion metric)
get_interspersed_kmers <- function(kmer, input_mx) {
  kmer.index <- which(colnames(input_mx)==kmer)
  a <- input_mx[kmer.index,which(input_mx[kmer.index,] > 0)]
  a.names <- names(which(input_mx[kmer.index,] > 0))
  b <- input_mx[which(input_mx[,kmer.index] > 0), kmer.index]
  b.names <- names(which(input_mx[,kmer.index] > 0))
  ab <- c(a,b)
  ab.names <- c(a.names, b.names)
  names(ab) <- ab.names
  return(ab)
}

# function to calculate GC content of a kmer
Calc_GC_content <- function (kmer_string) {
  string <- BString(kmer_string)
  gc.content <- c(letterFrequency(string, letters=c("CG"), as.prob=F)/(length(string)))
  return(gc.content)
}

AAAACCCT_1373 <- get_interspersed_kmers("AAAACCCT", interspersion.metric.matrix_1373)
hist(sapply(names(AAAACCCT_1373), Calc_GC_content), xlab="GC content", main="1373: GC of kmers interspersed with telomere")
length(AAAACCCT_1373)
AAAACCCT_1373


# these are the kmers that were found with Tandem Repeats Finder to be on the chloroplast assembly
cp_kmers <- c("AAAATAAAGTGT", "AAAATATATAAATATAGCT", "AAATACCTTACGGGAATAT")

# which kmers are interspersed with the telomere repeat in each of the ancestors?

AAAACCCT_1952 <- get_interspersed_kmers("AAAACCCT", interspersion.metric.matrix_1952)
AAAACCCT_1952
hist(sapply(names(AAAACCCT_1952), Calc_GC_content), xlab="GC content", main="1952: GC of kmers interspersed with telomere")
length(AAAACCCT_1952)

AAAACCCT_2342 <- get_interspersed_kmers("AAAACCCT", interspersion.metric.matrix_2342)
AAAACCCT_2342
hist(sapply(names(AAAACCCT_2342), Calc_GC_content), xlab="GC content", main="2342: GC of kmers interspersed with telomere")
length(AAAACCCT_2342)

AAAACCCT_2344 <- get_interspersed_kmers("AAAACCCT", interspersion.metric.matrix_2344)
AAAACCCT_2344
hist(sapply(names(AAAACCCT_2344), Calc_GC_content), xlab="GC content", main="2344: GC of kmers interspersed with telomere")
length(AAAACCCT_2344)

AAAACCCT_2931 <- get_interspersed_kmers("AAAACCCT", interspersion.metric.matrix_2931)
AAAACCCT_2931
length(AAAACCCT_2931)
hist(sapply(names(AAAACCCT_2931), Calc_GC_content), xlab="GC content", main="2931: GC of kmers interspersed with telomere")

AAAACCCT_2937 <- get_interspersed_kmers("AAAACCCT", interspersion.metric.matrix_2937)
AAAACCCT_2937
length(AAAACCCT_2937)
hist(sapply(names(AAAACCCT_2937), Calc_GC_content), xlab="GC content", main="2937: GC of kmers interspersed with telomere")

possible_cp_telomere <- c("AAATAACAAT", "AAATACCTTACGGGAATAT", "AAATACTTACGGGAATAT", "AAAATATATAAATATAGCT", "AAAGAAGTATATAAAT", "AAATACCTTACGGGAATAT", "AAATACTTACGGGAATAT", "AAATTTATATACTCC", "AATAGATAATAT", "AAATAGCAGTATAT", "AAAAAGGTAAATGTATTTAT", "AAAATAAAGTGT", "AAAATATATAAATATAGCT", "AAATACCTTACGGGAATAT", "AAATTTATATACTCC", "AAAATAAAGTGT", "AAATAGCAGTATAT", "AAAAAGGTAAATGTATTTAT", "AAATACCTTACGGGAATAT", "AAATACTTACGGGAATAT")

length(possible_cp_telomere)
cp_kmers %in% possible_cp_telomere

summary(as.factor(possible_cp_telomere))

```
Check interspersion of putative chloroplast kmers  

```{r}

AAATACCTTACGGGAATAT_1373 <- get_interspersed_kmers("AAATACCTTACGGGAATAT", interspersion.metric.matrix_1373)
AAATACCTTACGGGAATAT_1373

AAATACCTTACGGGAATAT_1952 <- get_interspersed_kmers("AAATACCTTACGGGAATAT", interspersion.metric.matrix_1952)
AAATACCTTACGGGAATAT_1952

AAATACCTTACGGGAATAT_2342 <- get_interspersed_kmers("AAATACCTTACGGGAATAT", interspersion.metric.matrix_2342)
AAATACCTTACGGGAATAT_2342

AAATACCTTACGGGAATAT_2344 <- get_interspersed_kmers("AAATACCTTACGGGAATAT", interspersion.metric.matrix_2344)
AAATACCTTACGGGAATAT_2344

AAATACCTTACGGGAATAT_2931 <- get_interspersed_kmers("AAATACCTTACGGGAATAT", interspersion.metric.matrix_2931)
AAATACCTTACGGGAATAT_2931

AAATACCTTACGGGAATAT_2937 <- get_interspersed_kmers("AAATACCTTACGGGAATAT", interspersion.metric.matrix_2937)
AAATACCTTACGGGAATAT_2937
length(AAATACCTTACGGGAATAT_2937)

```
## What is the interspersion patterns of the repeats that have interesting patterns of interspersion in the correlation plots? 
```{r}

(AGGC_1373 <- get_interspersed_kmers("AGGC", interspersion.metric.matrix_1373))
(AGGC_1952 <- get_interspersed_kmers("AGGC", interspersion.metric.matrix_1952))
(AGGC_2342 <- get_interspersed_kmers("AGGC", interspersion.metric.matrix_2342))
(AGGC_2344 <- get_interspersed_kmers("AGGC", interspersion.metric.matrix_2344))
(AGGC_2931 <- get_interspersed_kmers("AGGC", interspersion.metric.matrix_2931))
(AGGC_2937 <- get_interspersed_kmers("AGGC", interspersion.metric.matrix_2937))

(AGGCGC_1373 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_1373))
(AGGCGC_1952 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_1952))
(AGGCGC_2342 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2342))
(AGGCGC_2344 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2344))
(AGGCGC_2931 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2931))
(AGGCGC_2937 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2937))

(ACCAGC_1373 <- get_interspersed_kmers("ACCAGC", interspersion.metric.matrix_1373))
(ACCAGC_1952 <- get_interspersed_kmers("ACCAGC", interspersion.metric.matrix_1952))
(AGGCGC_2342 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2342))
(AGGCGC_2344 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2344))
(AGGCGC_2931 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2931))
(AGGCGC_2937 <- get_interspersed_kmers("AGGCGC", interspersion.metric.matrix_2937))

(ACAGGC_1373 <- get_interspersed_kmers("ACAGGC", interspersion.metric.matrix_1373))
(ACAGGC_1952 <- get_interspersed_kmers("ACAGGC", interspersion.metric.matrix_1952))
(ACAGGC_2342 <- get_interspersed_kmers("ACAGGC", interspersion.metric.matrix_2342))
(ACAGGC_2344 <- get_interspersed_kmers("ACAGGC", interspersion.metric.matrix_2344))
(ACAGGC_2931 <- get_interspersed_kmers("ACAGGC", interspersion.metric.matrix_2931))
(ACAGGC_2937 <- get_interspersed_kmers("ACAGGC", interspersion.metric.matrix_2937))

(AC_1373 <- get_interspersed_kmers("AC", interspersion.metric.matrix_1373))
length(AC_1373)
# all the kmers that are interspersed with others are interspersed with AC

```
# Get number of other kmers each kmer is interspersed with  

```{r}
# function returns number of kmers a kmer with a given index is interspersed with
get_total_interspersion <- function(kmer, input_mx ) {
  
  if (kmer.index == 1) { 
    return (sum(interspersion.metric.matrix[kmer.index, (kmer.index + 1):ncol(interspersion.metric.matrix) ]))
  } else if (kmer.index == ncol(interspersion.metric.matrix)) {
    return (sum(interspersion.metric.matrix[1:(kmer.index-1), kmer.index ]))
  } else {
    total_rows <- sum(interspersion.metric.matrix[kmer.index, (kmer.index + 1):ncol(interspersion.metric.matrix) ])
    total_cols <- sum(interspersion.metric.matrix[1:(kmer.index-1), kmer.index ])
    total <- total_rows + total_cols
    return (total)
  }
}

```
Look at the kmers that are interesting because their interspersion matrices - sometimes a positive correlation and sometimes negative. Make a bar chart par(mfrow)  

```{r}
par(mfrow=c(3,2))
barplot(AGGCGC_1373, names.arg=names(AGGCGC_1373), las=2, cex.names=0.7, cex.axis=0.8, main="AGGCGC_1373")
barplot(AGGCGC_1952, names.arg=names(AGGCGC_1952), las=2, cex.names=0.7, cex.axis=0.8, main="AGGCGC_1952")
barplot(AGGCGC_2342, names.arg=names(AGGCGC_2342), las=2, cex.names=0.7, cex.axis=0.8, main="AGGCGC_2342")
barplot(AGGCGC_2344, names.arg=names(AGGCGC_2344), las=2, cex.names=0.7, cex.axis=0.8, main="AGGCGC_2344")
barplot(AGGCGC_2931, names.arg=names(AGGCGC_2931), las=2, cex.names=0.7, cex.axis=0.8, main="AGGCGC_2931")
barplot(AGGCGC_2937, names.arg=names(AGGCGC_2937), las=2, cex.names=0.7, cex.axis=0.8, main="AGGCGC_2937")

barplot(ACAGGC_1373, names.arg=names(ACAGGC_1373), las=2, cex.names=0.7, cex.axis=0.8, main="ACAGGC_1373")
barplot(ACAGGC_1952, names.arg=names(ACAGGC_1952), las=2, cex.names=0.7, cex.axis=0.8, main="ACAGGC_1952")
barplot(ACAGGC_2342, names.arg=names(ACAGGC_2342), las=2, cex.names=0.7, cex.axis=0.8, main="ACAGGC_2342")
barplot(ACAGGC_2344, names.arg=names(ACAGGC_2344), las=2, cex.names=0.7, cex.axis=0.8, main="ACAGGC_2344")
barplot(ACAGGC_2931, names.arg=names(ACAGGC_2931), las=2, cex.names=0.7, cex.axis=0.8, main="ACAGGC_2931")
barplot(ACAGGC_2937, names.arg=names(ACAGGC_2937), las=2, cex.names=0.7, cex.axis=0.8, main="ACAGGC_2937")

library(ggplot2)
temp <- upper.tri(interspersion.metric.matrix_1373, diag = T)
# you already have triangular matrix
```
# Kmers that are related closely by sequence (see Figure 3 in manuscript) - do they have interesting patterns of interspersion?

```{r}
related_kmers_AG <- c("ACGC", "ACGCC", "ACGCCC", "AGGC", "AGGCGG", "AGGCGC","AGGGC",  "AGGGGC", "AGAGGC", "AGCGGC", "AGCGGG", "AGCCGC", "AGCGCC", "AGCCGG", "AGGGCC")

related_kmers_AC <- c("ACCGGC", "ACCCGC", "ACACGC", "ACAGGC", "ACAGGG", "ACCAGC", "ACCCGG", "ACTGGGC")

related_kmers_AG_1373 <- related_kmers_AG[related_kmers_AG %in% colnames(interspersion.metric.matrix_1373)] 
interspersion.metric.matrix_1373[related_kmers_AG_1373, related_kmers_AG_1373]
related_kmers_AC_1373 <- related_kmers_AC[related_kmers_AC %in% colnames(interspersion.metric.matrix_1373)]
interspersion.metric.matrix_1373[related_kmers_AC_1373, related_kmers_AC_1373]
# not much interspersion between related kmers

related_kmers_AG_1952 <- related_kmers_AG[related_kmers_AG %in% colnames(interspersion.metric.matrix_1952)] 
interspersion.metric.matrix_1952[related_kmers_AG_1952, related_kmers_AG_1952]
related_kmers_AC_1952 <- related_kmers_AC[related_kmers_AC %in% colnames(interspersion.metric.matrix_1952)]
interspersion.metric.matrix_1952[related_kmers_AC_1952, related_kmers_AC_1952]
# not much interspersion between related kmers

related_kmers_AG_2342 <- related_kmers_AG[related_kmers_AG %in% colnames(interspersion.metric.matrix_2342)] 
interspersion.metric.matrix_2342[related_kmers_AG_2342, related_kmers_AG_2342]
related_kmers_AC_2342 <- related_kmers_AC[related_kmers_AC %in% colnames(interspersion.metric.matrix_2342)]
interspersion.metric.matrix_2342[related_kmers_AC_2342, related_kmers_AC_2342]
# more interspersions in in the AG kmers, but not seeming to make pattern in accordance to the correlation matrix

related_kmers_AG_2344 <- related_kmers_AG[related_kmers_AG %in% colnames(interspersion.metric.matrix_2344)] 
interspersion.metric.matrix_2344[related_kmers_AG_2344, related_kmers_AG_2344]
related_kmers_AC_2344 <- related_kmers_AC[related_kmers_AC %in% colnames(interspersion.metric.matrix_2344)]
interspersion.metric.matrix_2344[related_kmers_AC_2344, related_kmers_AC_2344]
# all interspersed kmers are correlated positively or negatively

related_kmers_AG_2931 <- related_kmers_AG[related_kmers_AG %in% colnames(interspersion.metric.matrix_2931)] 
interspersion.metric.matrix_2931[related_kmers_AG_2931, related_kmers_AG_2931]
related_kmers_AC_2931 <- related_kmers_AC[related_kmers_AC %in% colnames(interspersion.metric.matrix_2931)]
interspersion.metric.matrix_2931[related_kmers_AC_2931, related_kmers_AC_2931]
# not many interspersed kmers

related_kmers_AG_2937 <- related_kmers_AG[related_kmers_AG %in% colnames(interspersion.metric.matrix_2937)] 
interspersion.metric.matrix_2937[related_kmers_AG_2937, related_kmers_AG_2937]
related_kmers_AC_2937 <- related_kmers_AC[related_kmers_AC %in% colnames(interspersion.metric.matrix_2937)]
interspersion.metric.matrix_2937[related_kmers_AC_2937, related_kmers_AC_2937]


```
# Kmers that are in the correlation matrix - what are their intersperion patterns?  

```{r}
corr_mx_kmers <- c("C", "AC", "ACC", "AAC", "ATC", "AGC", "AGG", "AAT", "CCG", "AGCGGC", "AGCGCC", "AGCGGG", "AGCCGC", "AGCCGG", "AGGGGC", "AGGGC", "AGGCGC", "AGGC", "AGGCGG", "ACGC", "ACGCC", "ACGCCC", "ACCCGC", "ACCAGC", "ACCGGC", "ACAGGC", "ACAGGG", "ACACGC", "AAAGT")
length(corr_mx_kmers)
#write.csv(corr_mx_kmers, file="~/Documents/AndyLab/Chlamy/R_Analysis/corr_mx_kmers.csv", quote = F, row.names = F)

corr_mx_kmers_1373 <- corr_mx_kmers[corr_mx_kmers %in% colnames(interspersion.metric.matrix_1373)]
corrkmers_1373 <- interspersion.metric.matrix_1373[corr_mx_kmers_1373, corr_mx_kmers_1373]

get_corrkmer_mx <- function(interspersion_mx, Ancestor) {
  outdir <- "~/Documents/AndyLab/Chlamy/R_Analysis/data_out/"
  filename <- paste0(outdir, Ancestor, "_int.summary", ".csv")
  kmers_to_incude <- corr_mx_kmers[corr_mx_kmers %in% colnames(interspersion_mx)] 
  output_mx <- interspersion_mx[kmers_to_incude, kmers_to_incude]
  write.csv(output_mx, file=filename, quote=F)
}
get_corrkmer_mx(interspersion.metric.matrix_1373, "CC-1373")
get_corrkmer_mx(interspersion.metric.matrix_1952, "CC-1952")
get_corrkmer_mx(interspersion.metric.matrix_2342, "CC-2342")
get_corrkmer_mx(interspersion.metric.matrix_2344, "CC-2344")
get_corrkmer_mx(interspersion.metric.matrix_2931, "CC-2931")
get_corrkmer_mx(interspersion.metric.matrix_2937, "CC-2937")

```
**All of the common kmers were interspersed to some extent in at least one ancestor.**  

```{r}

all_interspersed_kmers <- Reduce(union, list(colnames(interspersion.metric.matrix_1373), colnames(interspersion.metric.matrix_1952), colnames(interspersion.metric.matrix_2342), colnames(interspersion.metric.matrix_2344), colnames(interspersion.metric.matrix_2931), colnames(interspersion.metric.matrix_2937)))

all_interspersed_kmers %in% common.kmer.union
common.kmer.union %in% all_interspersed_kmers                             
length(colnames(interspersion.metric.matrix_1373))
ncol((interspersion.metric.matrix_1952))
